---
title: "Introduction to cpnr"
author: Gregor Mathes
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to cpnr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(cpnr)
```

# cpnr: Tutorials for conservation paleobiology  
  
The [Conservation Paleobiology Network (CPN)](https://conservationpaleorcn.org/) aims to transform conservation paleobiology into an applied science that informs conservation and restoration efforts. The project is envisioned as a grass-roots effort to develop strategies for communicating geohistorical data to stakeholder groups.  
  
To facilitate interdisciplinary conservation work, the CPN now offers tutorials. These will cover various topics starting at an introductory level and aim to show students and scientists how to process geohistorical data. This document shows how to use the `cpnr` package and how to start a tutorial. 
