---
title: "Quick introduction to cpnr"
author: Gregor Mathes
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick introduction to cpnr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(cpnr)
```

# cpnr: Interactive tutorials for conservation paleobiology  
  
This is a quick tour through the `cpnr` package amd will show how you how to start the *divdyn* tutorial.
  
### How to start  
  
After installing and loading the package, you can see what tutorials are currently available with  the `show_tutorials` function:  

```{r show_tutorials}
show_tutorials()
```  
  
We can see that there is a tutorial called *divdyn*. Let's get a bit more detail for this one.  
  
### Further details  
  
If the short description returned from `show_tutorials` is not sufficient for you to see what the tutorial contains, you can access a more detailed description via `show_description`. We need the tutorial name as an argument for the function:  
  
```{r show_description}
show_description(tutorial = "divdyn")
```   
  
If that looks good to you, let's start the tutorial. 
  
### What you need  
  
Each tutorial relies on various packages that might not be installed on your device. Running the tutorial without these packages will fail. It is therefore necessary to check which packages are needed to run the tutorial, and to install them if necessary.  
  
```{r show_dependencies}
show_dependencies(tutorial = "divdyn")
```  
  
To load all these dependencies needed to run a tutorial, and to install the missing ones, you can run `load_depencies`:  
  
```{r load_dependencies, eval=FALSE}
# get dependencies
packages <- show_dependencies(tutorial = "divdyn")

# load or install them
load_dependencies(packages = packages)
```  
  
### Access the tutorial  
  
Now you should know the name of the tutorial you want to run and what it does based on the description. Further, you have loaded and installed all dependencies needed to run the tutorial. Now you are ready to run it. The `start_tutorial` function will initiate the interactive tutorial and open it in a browser:  
  
```{r start_tutorial, eval=FALSE}
start_tutorial(tutorial = "divdyn")
```  
  
Make sure to end the tutorial with the stop button in the R console when you are done.    
  




  



